---
author : "Clementine"
date : "2023-02-28"
title : "Project overview"
output: html_document
---

# Questions

How did recruitment dynamics changed in the Quebec boreal forest between the 
periods 1970-1981 and 2005-2018 ?

   - In term of presence and density
   - For the common species
   - How did perturbation influenced those changes
   - What is the effect of other parameters (biotic or abiotic)

*further question :*

   - *Growth from saplings to trees*
   - *What does it means for the future of the boreal forest ?*
   - *Ecosystem shifts ?*

To keep in mind for my report : **Ecosystem functionning**

# Method

## Data

All the data from saplings
Biologic parameters from tree_data
Perturbation from pep_ori

- 2 time periodes : 1970-1981 and 2005-2018 (to be confirmed)
- Species : Sapin baumier, erable rouge, erable à sucre, bouleau jaune, 
    bouleau à papier, epinette noire (beginning with sapin baumier and erable 
    rouge to make the model)
- Zone : All (to be confirmed)
- Other selections :
   - ph, epmatorg et cl_drai connues (au moins une fois pour la placette)
    - pas de plantation
    - date de perturbation connues

*Would be interesting to think about including  seedlings*

## Model

Indépendant pour chaque espèces

### Parameters

### Abiotic

- ph_humus
- organic matter thickness
- soil drainage ?
- topographique ? (pente, altitude, exposition)

#### Biotic

- Basal area of the Species
- Total basal area
- Couvert ? (I don't know if I have enough data)
- Species richness could be interesting

Perhaps there are better ways to represent the adult population

### Geographic

- Latitude
- Longitude

### Perturbation

- Type of disturbance
- Time since disturbance

### Effect of the periode

On :

- the intercept
- the coef for latitude

I assume that the effect of botic and abiotic parameters on recruitment are the 
same for the 2 periodes (no evolution)

## Model

Two part hurdle model :
    - First part : presence or absence of the species saplings
    $$PA_i = \mathcal{B}(p_i)$$
    - Second part : density of saplings
    $$NB_i = \mathcal{P}(\lambda_i * presence_i)$$

Avec (a completer avec les autres paramètres qui m'interessent):\
\

$$logit(p_i) = Intercept_{t} + \alpha_{t} * latitude_i +
\gamma_{disturbance_i} + \gamma'_{disturbance_i} * TSD_i +
\beta_1 * ph_i + \beta_2 * epmatorg_i + \beta_4 * longitude_i +
\delta_{cl.drai_i}$$

$$log(\lambda_i) = Intercept_{t} + \alpha_{t} * latitude_i +
\gamma_{disturbance_i} + \gamma'_{disturbance_i} * TSD_i +
\beta_1 * ph_i + \beta_2 * epmatorg_i + \beta_4 * longitude_i +
\delta_{cl.drai_i}$$

\
__Note__ : different coef for lambda and p

In consideration :

- There might be parameters to add to the square knowing that the relation is
not linear (example ph, TSD)
- Living things don't only have linear or even polynomial relations, there are
also limits (of ph for example simply for survival)
- Interaction between parameters

# Results exemple

With reduced data for the exemple :

- random sample of 1000 lines
- For ERR (erable rouge)
- 1 type of disturbance only : logging (all other pe are concidered as not 
disturbed)

__Note__ : I need to truncate the Poisson law but I have a problem with it

```{r, include=FALSE, echo=FALSE}
source("data.R")
```

## Model code

```{r}
model_pert <- function(){
    # likelihood
    for(i in 1:N){
        pa[i] ~ dbern(p[i])
        nb[i] ~ dpois(pa[i] * lambda[i]) # T(0,)

        logit(p[i]) <-
            intercept_pa[t[i]] +
            lat_pa[t[i]] * latitude[i] +
            lon_pa * longitude[i] +
            ph_pa * ph_humus [i] +
            ph_pa2 * (ph_humus [i])**2 +
            epm_pa * epmatorg[i] +
            beta * is_logging[i] +
            gamma * TSlogging[i]
        log(lambda[i]) <- 
            intercept_nb[t[i]] +
            lat_nb[t[i]] * latitude[i] +
            lon_nb * longitude[i] +
            ph_nb * ph_humus [i] +
            ph_nb2 * (ph_humus [i])**2 +
            epm_nb * epmatorg[i]
        }

    # priors
    for (t in 1:2){
        intercept_pa[t] ~ dunif(-10, 10)
        lat_pa[t] ~ dunif(-10, 10)
        intercept_nb[t] ~ dunif(-10, 10)
        lat_nb[t] ~ dunif(-10, 10)
    }
    lon_pa ~ dunif(-10, 10)
    ph_pa ~ dunif(-10, 10)
    epm_pa ~ dunif(-10, 10)
    lon_nb ~ dunif(-10, 10)
    ph_nb ~ dunif(-10, 10)
    epm_nb ~ dunif(-10, 10)
    ph_nb2 ~ dunif(-10, 10)
    ph_pa2 ~ dunif(-10, 10)
    beta ~ dunif(-10, 10)
    gamma ~ dunif(-1, 1)
}
```

```{r, include=FALSE, echo=FALSE}

full_data1_err <- full_data1 %>%
    filter(essence == "ERR")

full_data1_err <- full_data1_err %>%
    mutate(periode = ifelse(year_measured < 1982, 1,
        ifelse(year_measured > 2000, 2, NA))) %>%
        select(periode, latitude, longitude, ph_humus, epmatorg, pourcpierr,
            presence_gaule, all_cl,logging) %>%
        na.omit() %>%
        mutate(is_logging = ifelse(logging > 0, 1, 0))

full_data1_err <- full_data1_err %>%
    mutate(
        latitude = scale(latitude)[,1],
        longitude = scale(longitude)[,1],
        ph_humus = scale(ph_humus)[,1],
        epmatorg = scale(epmatorg)[,1],
    )
print(paste("Il y a ", nrow(full_data1_err), "lignes pour l'essence ERR"))

# 1000 lignes au hasard
netud = 1000
full_data1_err <- full_data1_err %>%
    sample_n(netud)
print(paste("J'étudie",nrow(full_data1_err), "lignes pour commencer"))
```

```{r, include=FALSE, echo=FALSE}
jags_data <- list(
    N = nrow(full_data1_err),
    t = full_data1_err$periode,
    latitude = full_data1_err$latitude,
    longitude = full_data1_err$longitude,
    ph_humus = full_data1_err$ph_humus,
    epmatorg = full_data1_err$epmatorg,
    pa = full_data1_err$presence_gaule,
    nb = full_data1_err$all_cl,
    is_logging = full_data1_err$is_logging,
    TSlogging = full_data1_err$logging
)
```
    
```{r, include=FALSE, echo=FALSE}
old <- Sys.time()

out_model <- jags(
    model.file = model_pert,
    data = jags_data,
    n.chains = 3,
    n.iter = 10000,
    inits = function(){
        list(
            intercept_pa = runif(2, -10, 10),
            lat_pa = runif(2, -10, 10),
            intercept_nb = runif(2, -10, 10),
            lat_nb = runif(2, -10, 10),
            lon_pa = runif(1, -1, 1),
            ph_pa = runif(1, -1, 1),
            epm_pa = runif(1, -1, 0),
            lon_nb = runif(1, -1, 1),
            ph_nb = runif(1, -1, 1),
            epm_nb = runif(1, -1, 1),
            ph_pa2 = runif(1, -1, 1),
            ph_nb2 = 0.001,
            beta = 2.45,
            gamma = -0.07
        )},
    parameters.to.save = c("intercept_pa", "lat_pa", "intercept_nb", "lat_nb",
                           "lon_pa", "ph_pa", "epm_pa",
                           "lon_nb", "ph_nb", "epm_nb",
                           "ph_pa2", "ph_nb2", "beta", "gamma"),
    n.burnin = 1000,
    n.thin = 1,
    DIC = TRUE
)

runtime = Sys.time() - old
```

## Model output

```{r}
print(runtime)
```

```{r}
print(out_model)
```

## Parameters output

```{r }
mcmc_intervals(as.mcmc(out_model),
    pars = c("intercept_pa[1]", "intercept_pa[2]", "lat_pa[1]", "lat_pa[2]",
        "intercept_nb[1]", "intercept_nb[2]", "lat_nb[1]", "lat_nb[2]",
        "lon_pa", "ph_pa", "epm_pa", "lon_nb", "ph_nb", "epm_nb",
        "ph_pa2", "ph_nb2", "beta", "gamma"))
```

__Note__ : autoceralation and convergence are great

__Note__ : priors may not be very good

```{r}
traceplot(out_model, ask = FALSE)
```

```{r}
autocorr.plot(out_model, ask = FALSE)
```

